---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(fpp3)
library(rsample)
```

```{r}
global_economy

global_economy |> 
  filter(Country == "Australia") |>
  autoplot(GDP/Population) +
  labs(title= "GDP per capita", y = "$US")
```

\`\`\`{r}library(fpp3)}

```{r}
library(fpp3)
file_name= "db172003-6497-4adc-a4e9-cfbf12022d28"
source("loading_data.R")
train_path = "data/"
path = paste0(train_path,file_name,".result")
data <- read_data(path, 1)

```

------------------------------------------------------------------------

## We create a dataframe for timeseries with our data.

---
Now we read form the CSV f the rows with the SoH
---

```{r}
csv_path = paste0(train_path, "experiment_results.csv")
soh_data = get_soh(csv_path, file_name)
soh_data = soh_data / soh_data[1]
```

We first filter the first cycle and the discharge step and see that the two of then behave in a similar way.

We can then first try to predict the behavior of the next voltage just using the voltarge itself.

```         
```

```{r}
get_measurement <- function(data, current_cycle) {
  measurement <-  data |>
  filter(cycle ==current_cycle, Step == 5) %>%
  select(c(C,V, cycle,SoC), time) %>%  # Selecting the variable 'V' and 'time'
  index_by(time_10 = floor_date(time, "10 seconds")) %>%  # Resample every 10 seconds
  summarize(across(everything(), \(x) mean(x, na.rm = TRUE))) %>%  # Aggregate mean values
  fill_gaps()
  measurement
}
get_measurement_up <- function(data, current_cycle) {
  measurement <-  data |>
  filter(cycle <=current_cycle, Step == 5) %>%
  select(c(C,V,cycle, SoC), time) %>%  # Selecting the variable 'V' and 'time'
  index_by(time_10 = floor_date(time, "10 seconds")) %>%  # Resample every 10 seconds
  summarize(across(everything(), \(x) mean(x, na.rm = TRUE))) %>%  # Aggregate mean values
   fill_gaps() 
  measurement
}
get_measurement_complete_up <- function(data, current_cycle) {
  measurement <-  data |>
  filter(cycle <=current_cycle) %>%
  select(c(C,V, SoC, Step, cycle), time) %>%  # Selecting the variable 'V' and 'time'
  index_by(time_10 = floor_date(time, "10 seconds")) %>%  # Resample every 10 seconds
  summarize(across(everything(),\(x) mean(x, na.rm = TRUE))) %>%  # Aggregate mean values
   fill_gaps() 
  measurement
}
measurement <- get_measurement(data, 1)
```

```{r}
autoplot(measurement, .vars = c(V), color = "blue") +   
  labs(title = "Charge-Discharge Step Discharge Profile Voltage",
       x = "Total Hours",
       y = "Voltage") +
  scale_color_manual(values = "blue")

```

```{r}
autoplot(measurement, .vars = c(C), color = "red") +   
  labs(title = "Charge-Discharge Step Discharge Profile Current",
       x = "Total Hours",
       y = "Current")
```

```{r}
voltage <- measurement %>% select(V,time)
```

```{r}
train_prop = 0.8

# Get the total number of rows
n <- nrow(voltage)

# Calculate the index for 80%
train_size <- floor(train_prop * n)

# Split the data
train_data <- voltage %>%
  slice(1:train_size)  # First 80% for training

test_data <- voltage %>%
  slice((train_size + 1):n)  # Remaining 20% for testing
```

```{r}
# Fit the models
voltage_fit <- train_data|> fill_gaps() |>
  model(
    Mean = MEAN(V),
    `Naïve` = NAIVE(V),
  )
```

```{r}
voltage_fc <- voltage_fit |> forecast(h = nrow(test_data))
# Plot forecasts against actual values
voltage_fc |>
  autoplot(train_data, level = NULL) +
  autolayer(test_data, colour = "grey") + 
  labs(
    y = "Megalitres",
    title = "Forecasts for quarterly beer production"
  ) +
  guides(colour = guide_legend(title = "Forecast"))


```

### Residuals

The "residuals" in a time series model are what is left over after fitting a model. The residuals are equal to the difference between the observations and the corresponding fitted values: et=yt−\^yt.

We call these "**innovation residuals**". For example, suppose we modelled the logarithms of the data, wt=log(yt). Then the innovation residuals are given by wt−\^wt whereas the regular residuals are given by yt−\^yt

```{r}
aug <- train_data|> fill_gaps() |>
  model(NAIVE(V)) |> augment()
autoplot(aug, .innov) +
  labs(y = "Voltage",
       title = "Residuals from the MEAN method")
```

```{r}
aug |>
  ggplot(aes(x = .innov)) +
  geom_histogram() +
  labs(title =  "Histogram of residuals")
```

```{r}
aug |>
  ACF(.innov) |>
  autoplot() +
  labs(title = "Residuals from the MEAN method")
```

```{r}
train_data|> fill_gaps() |>
  model(NAIVE(V)) |>
  gg_tsresiduals()

```

### Forecast distributions

### Prediction intervals

### One-step prediction intervals & Multi-step prediction intervals

```{r}
voltage |>
  model(NAIVE(V)) |>
  forecast(h = 10) |>
  hilo()
```

```{r}
voltage |>
  model(NAIVE(V)) |>
  forecast(h = 12) |>
  autoplot(voltage) +
  labs(title="Voltage of the cycle", y="V" )
```

```         
Bootstrapping 
```

```{r}
fit <- voltage |>
  model(NAIVE(V))
sim <- fit |> generate(h = 100, times = 5, bootstrap = TRUE)
sim
```

```{r}
# Calculate the middle point of the dataset
middle_index <- floor(nrow(voltage) / 2)
voltage%>%slice(middle_index:n()) |> 
  ggplot(aes(x = time_10)) +
  geom_line(aes(y = V)) +
  geom_line(aes(y = .sim, colour = as.factor(.rep)),
    data = sim) +
  labs(title="Voltage prediction with 5 path using NAIVE METHOD", y="V" ) +
  guides(colour = "none")
```

```{r}
fit <- train_data |>
  model(
    Mean = MEAN(V),
    `Naïve` = NAIVE(V),
    `Seasonal naïve` = SNAIVE(V),
    Drift = RW(V ~ drift())
  )

data_fc <- fit |>
  forecast(h = nrow(test_data))
middle_index <- floor(nrow(train_data) / 2)
plot_train_data <- train_data%>%slice(middle_index:n())
data_fc |>
  autoplot(
    plot_train_data,
    level = NULL
  ) +autolayer(test_data, colour = "grey") +
  labs(
    y = "Voltage",
    title = "Battery data"
  ) +
  guides(colour = guide_legend(title = "Forecast"))
```

```{r}
accuracy(data_fc, test_data)
```

```{r}
measurement |>
  pivot_longer(c(C, V), names_to="Series") |>
  autoplot(value) +
  labs(y = "Battery Measurement")
```

```{r}
measurement |>
  ggplot(aes(x = C, y = V)) +
  labs(y = "Voltage (V)",
       x = "Current (C)") +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
measurement |>  model(TSLM(V ~ C)) |>
  report()

```

SoC

### Multiple linear regression

```{r}
measurement_soc <- measurement %>% select(-time)
```

```{r}
measurement_soc |>
  pivot_longer(-time_10) |>
  ggplot(aes(time_10, value, colour = name)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  guides(colour = "none") +
  labs(y="% change")
```

## Least squares estimation

```{r}
fit_V <- measurement_soc |>
  model(tslm = TSLM(V ~ C))
report(fit_V)

```

```{r}
augment(fit_V) |>
  ggplot(aes(x = time_10)) +
  geom_line(aes(y = V, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  labs(y = NULL,
    title = "SoC estimation using V and C"
  ) +
  scale_colour_manual(values=c(Data="black",Fitted="#D55E00")) +
  guides(colour = guide_legend(title = NULL))
```

### Trend

```{r}
fit_V_trend <- measurement |>
  model(tslm = TSLM(V ~ C + trend()))
report(fit_V_trend)
```

```{r}
augment(fit_V_trend) |>
  ggplot(aes(x = time_10)) +
  geom_line(aes(y = V, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  labs(y = NULL,
    title = "Voltage estimatin usign Current and trend"
  ) +
  scale_colour_manual(values=c(Data="black",Fitted="#D55E00")) +
  guides(colour = guide_legend(title = NULL))
```

```{r}
fit_SoC <- measurement |>
  model(tslm = TSLM(SoC ~ V + C))
report(fit_SoC)
```

```         
```

```{r}
augment(fit_SoC) |>
  ggplot(aes(x = time_10)) +
  geom_line(aes(y = SoC, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  labs(y = NULL,
    title = "SoC estimation using V and C"
  ) +
  scale_colour_manual(values=c(Data="black",Fitted="#D55E00")) +
  guides(colour = guide_legend(title = NULL))

```

```{r}
augment(fit_SoC) |>
  ggplot(aes(x = SoC, y = .fitted)) +
  geom_point() +
  labs(
    y = "Fitted (predicted values)",
    x = "Data (actual values)",
    title = "State of Charge of the battery"
  ) +
    scale_colour_manual(values=c("black","#D55E00")) +
  geom_abline(intercept = 0, slope = 1)
```

## Forecasting with regression

**Ex-ante forecasts**

**Ex-post forecasts**

```{r}

# Calculate the total number of unique hours in the dataset
total_hours <- n_distinct(floor_date(measurement$time_10, "hour"))

# Calculate the cutoff for the first 80% of the hours
cutoff_hour <- floor(0.8 * total_hours)

recent_measuremnet <- measurement%>% slice((floor(0.8 * n()) + 1):n())
past_measuremnet <- measurement%>% slice(0:(floor(0.8 * n()) ))
fit_SoC <- past_measuremnet |>
  model(
    lm = TSLM(SoC ~ C + V )
  )
future_scenarios <- scenarios(
  Increase = new_data(past_measuremnet,n= nrow(recent_measuremnet)) %>% mutate(C = recent_measuremnet$C, V = recent_measuremnet$V),
  names_to = "Scenario")

Increase = new_data(past_measuremnet,n= nrow(recent_measuremnet)) %>% mutate(C = recent_measuremnet$C, V = recent_measuremnet$V)

fc <- forecast(fit_SoC, new_data = Increase)

forecast(fit_SoC, new_data = Increase)|>
  autoplot(recent_measuremnet) +
  labs(title = "SoC", y = "% change")
accuracy(fc, recent_measuremnet)
```

```{r}

# Calculate the total number of unique hours in the dataset
total_hours <- n_distinct(floor_date(measurement$time_10, "hour"))

# Calculate the cutoff for the first 80% of the hours
cutoff_hour <- floor(0.8 * total_hours)

recent_measuremnet <- measurement%>% slice((floor(0.8 * n()) + 1):n())
past_measuremnet <- measurement%>% slice(0:(floor(0.8 * n()) ))
fit_SoC <- past_measuremnet |>
  model(
    lm = TSLM(SoC ~ C + V + trend())
  )
future_scenarios <- scenarios(
  Increase = new_data(past_measuremnet,n= nrow(recent_measuremnet)) %>% mutate(C = recent_measuremnet$C, V = recent_measuremnet$V),
  names_to = "Scenario")

Increase = new_data(past_measuremnet,n= nrow(recent_measuremnet)) %>% mutate(C = recent_measuremnet$C, V = recent_measuremnet$V)

fc <- forecast(fit_SoC, new_data = Increase)

forecast(fit_SoC, new_data = Increase)|>
  autoplot(recent_measuremnet) +
  labs(title = "SoC", y = "% change")
accuracy(fc, recent_measuremnet)
past_measuremnet  %>% model(
    lm = TSLM(SoC ~ C + V + trend())
  )|>
  gg_tsresiduals()
```

```{r}
report(fit_SoC)
```

### Correlation is not causation

It is important not to confuse correlation with causation, or causation with forecasting. A variable x may be useful for forecasting a variable y, but that does not mean x is causing y.

### Forecasting with correlated predictors

### Multicollinearity and forecasting

## Exponential Smoothing

```{r}
fit_ETS_SoC <- measurement |>
  model(ETS(SoC))
report(fit_ETS_SoC)
```

```{r}
fit_ETS_SoC <- measurement |>
  model(ETS(SoC ~ trend("A"))) 
report(fit_ETS_SoC)
```

```{r}
# Calculate the middle point of the dataset
middle_index <- floor(nrow(measurement) / 2)
measurement |>
  model(
    `No Trend` = ETS(SoC),
    `Trend` = ETS(SoC ~ error("A") +
                       trend("Ad", phi = 0.9) + season("N"))
  ) |>
  forecast(h = 50) |>
  autoplot(measurement %>%slice(middle_index:n()), level = NULL) +
  labs(title = "Australian population",
       y = "Millions") +
  guides(colour = guide_legend(title = "Forecast"))
```

## ARIMA

```{r}
library(urca)
adf_test <- ur.df(measurement$SoC, type = "drift")
summary(adf_test)

```

```{r}
caf_fit <- measurement |>
  model(arima210 = ARIMA(SoC ~ pdq(2,1,0)),
        arima013 = ARIMA(SoC ~ pdq(0,1,3)),
        stepwise = ARIMA(SoC),
        search = ARIMA(SoC, stepwise=FALSE))

```

```{r}
caf_fit |>
  forecast(h=100) |>
  filter(.model=='search') |>
  autoplot(measurement%>%slice(middle_index:n()))
```

`{glance(caf_fit) |> arrange(AICc) |> select(.model:BIC)}`

## Regression with ARIMA

```{r}
measurement |>
  pivot_longer(c(C, V, SoC),
               names_to = "var", values_to = "value") |>
  ggplot(aes(x = time, y = value)) +
  geom_line() +
  facet_grid(vars(var), scales = "free_y") +
  labs(title = "Battery cycling",
       y = "Signal Mesurement")
```

```{r}
# Calculate the total number of unique hours in the dataset
total_hours <- n_distinct(floor_date(measurement$time_10, "hour"))

# Calculate the cutoff for the first 80% of the hours
cutoff_hour <- floor(0.8 * total_hours)

recent_measuremnet <- measurement%>% slice((floor(0.8 * n()) + 1):n())
past_measuremnet <- measurement%>% slice(0:(floor(0.8 * n()) ))

fit <- past_measuremnet |>
  model(ARIMA(SoC ~ C  +  V + trend()))
futures <- new_data(past_measuremnet, n = nrow(recent_measuremnet)) %>%
  mutate(V = recent_measuremnet$V, C = recent_measuremnet$C)
fc_arima <- forecast(fit, new_data = futures) 
fc_arima |>
  autoplot(recent_measuremnet) +
  labs(y = "Percentage change")
accuracy(fc_arima, recent_measuremnet)
```

```{r}
fit <- past_measuremnet |>
  # Restrict data so models use same fitting period
  mutate(SoC = c(NA, NA, NA, SoC[4:nrow(past_measuremnet)])) |>
  # Estimate models
  model(
    lag0 = ARIMA(SoC ~ C  +  V + trend()),
    lag1 = ARIMA(SoC ~ C  +  V + lag(C) + lag(V)),
    lag2 = ARIMA(SoC ~ C  +  V + trend() + lag(C) + lag(V)+ lag(C, 2) + lag(V, 2)))
glance(fit)
```

```{r}
fit_best <- past_measuremnet |>
  model(ARIMA(SoC ~ C  +  V + lag(C) + lag(V)+ lag(C, 2) + lag(V, 2)))
report(fit_best)
```

```{r}
futures <- new_data(past_measuremnet, n = nrow(recent_measuremnet)) %>%
  mutate(V = recent_measuremnet$V, C = recent_measuremnet$C)
fc_arima <- forecast(fit_best, new_data = futures) 
fc_arima |>
  autoplot(recent_measuremnet) +
  labs(y = "Percentage change")
accuracy(fc_arima, recent_measuremnet)
```

```{r}
first_cycle <- get_measurement(data,1)
```

```{r}
autoplot(first_cycle, V)
```

```{r}

get_past_recent <- function(measurement, prop)
{
    # Calculate the total number of unique hours in the dataset
  total_hours <- n_distinct(floor_date(measurement$time, "hour"))
  
  # Calculate the cutoff for the first 80% of the hours
  cutoff_hour <- floor(prop * total_hours)
  
  recent_measuremnet <- measurement%>% slice((floor(prop * n()) + 1):n())
  past_measuremnet <- measurement%>% slice(0:(floor(prop * n()) ))
  return(list(recent_measuremnet = recent_measuremnet, past_measuremnet = past_measuremnet))
}


get_perform <- function(past_measuremnet, recent_measuremnet)
{
  models_fc <- past_measuremnet |> model(
    lm = TSLM(SoC ~ C + V + trend()))
  
  futures <- new_data(past_measuremnet, n = nrow(recent_measuremnet)) %>%
    mutate(V = recent_measuremnet$V, C = recent_measuremnet$C)
  fc_arima <- forecast(models_fc, new_data = futures) 
  # Create the plot
  forecast_plot <- fc_arima |>
    autoplot(recent_measuremnet) +
    labs(y = "Percentage change")
  accuracy_result <- accuracy(fc_arima, recent_measuremnet)
  # Return both the plot and the accuracy result as a list
  return(list(plot = forecast_plot, accuracy = accuracy_result))
}

```

```{r}
next_cycle = 1
measurement_cycle <- get_measurement(data = data,current_cycle = next_cycle)
past_recent <- get_past_recent(measurement_cycle, 0.9)
get_perform(past_recent$past_measuremnet, past_recent$recent_measuremnet)
```

What happen if we add all the regressos including step and Cycle

```{r}

get_perform_all <- function(past_measuremnet, recent_measuremnet)
{
  models_fc <- past_measuremnet |> model(
    lm = TSLM(SoC ~ C + V+ Step + cycle  + trend()))
  
  futures <- new_data(past_measuremnet, n = nrow(recent_measuremnet)) %>%
    mutate(V = recent_measuremnet$V, C = recent_measuremnet$C, Step = recent_measuremnet$Step, cycle = recent_measuremnet$cycle)
  fc_arima <- forecast(models_fc, new_data = futures) 
  # Create the plot
  forecast_plot <- fc_arima |>
    autoplot(recent_measuremnet) +
    labs(y = "Percentage change")
  accuracy_result <- accuracy(fc_arima, recent_measuremnet)
  # Return both the plot and the accuracy result as a list
  return(list(plot = forecast_plot, accuracy = accuracy_result))
}

next_cycle = 3
measurement_cycle <- get_measurement_complete_up(data = data,current_cycle = next_cycle)
past_recent <- get_past_recent_last_cycle(measurement_cycle)

```

```{r}

get_perform_all(past_recent$past_measurement, past_recent$recent_measurement)
```

ARIMA Methods are not valid to capture the behaviour of our signal SoC and hteri relationmship with previous step while the regrerssion modles can estimate teh SOC without a big error.\
It would be interesting to understand if the model has the same error when the degradation start to happen, for this we are going to select the last 10% of the discharge curve and compare the result of cycle 10 and 250.

```{r}
prop = 0.9
next_cycle = 10
measurement_cycle <- get_measurement(data = data,current_cycle = next_cycle)
past_recent <- get_past_recent(measurement_cycle, prop)
result_10 <- get_perform(past_recent$past_measuremnet, past_recent$recent_measuremnet)
result_10$accuracy
```

```{r}
next_cycle = 230
measurement_cycle <- get_measurement(data = data,current_cycle = next_cycle)
past_recent <- get_past_recent(measurement_cycle, prop)
result_250 <- get_perform(past_recent$past_measuremnet, past_recent$recent_measuremnet)
result_250$accuracy
```

```{r}
results_table <- data.frame(lm = numeric())
for (cycle in seq(1, 230, by = 5)) {
  measurement_cycle <- get_measurement(data = data,current_cycle = cycle)
  past_recent <- get_past_recent(measurement_cycle, prop)
  result <- get_perform(past_recent$past_measuremnet, past_recent$recent_measuremnet)
  RMSE = result$accuracy$RMSE
  # Agregamos los resultados a la tabla
  results_table <- rbind(results_table, data.frame(lm = RMSE[1]))
  print(paste("cycle : ",cycle) )
}
```

```{r}

# Convert the data to long format for faceting
df = tibble(results_table)
# Add an index column to the dataframe
df <- df %>%
  mutate(Index = row_number())

# Convert the data to long format for faceting
df_long <- df %>%
  pivot_longer(cols = -Index, names_to = "Model", values_to = "Value")

# Create the faceted plot
ggplot(df_long, aes(x = Index, y = Value)) +
  geom_line() +
  facet_wrap(~Model, scales = "free_y") +
  labs(x = "Index", y = "Value", title = "Model Comparison") +
  theme_minimal()

```
